Stack Safety is a Security Property

### 2021-04-16 tex/main.tex
Cheri related work: lighter, a little shorter

\paragraph{Formal Stack Safety on Capability Machines}
%
To our knowledge, the only line of work to date that has attempted
a positive and formal characterization of
stack safety is \citet{Skorstengaard+19b}, who introduce a calling
convention that uses local capabilities to preserve local state
encapsulation and well-bracketed control flow, using a logical
relation to reason about the stack safety of concrete
programs. Although the required hardware support is readily available
in capability machines like Cheri, this technique incurs significant
costs, because it requires the entire unused part of the stack to be
cleared whenever a security boundary is crossed. Their logical
relation captures capability safety without ``externally observable
side-effects (like console output or memory access traces)''
and can be used to reason about individual programs. \bcp{Not very clear:}In the
discussion, ``while [the authors] claim that [their] calling
convention enforces control-flow correctness, [they] do not prove a
general theorem that shows this, because it is not clear what such a
theorem should look like,'' noting that the correctness property
enjoyed by their technique ``is not made very explicit.''

StkTokens \citep{Skorstengaard+19} continues this line of work. Like
the earlier paper, it aims to protect the stack by enforcing local state
encapsulation and well-bracketed control flow. It does so by defining a new
calling convention that makes use of linear capabilities for stack and return
pointers. The convention operates on a single shared stack and requires that
protected components avoid compromising their own security by following certain
simple rules---like not leaking their private capability seals. More
precisely but still informally, local state encapsulation is defined as
restricting accesses to the range of memory allocated to the current stack
frame, and well-bracketed control flow as only allowing returns from the topmost
frame to the immediately adjacent frame below. Formally, it improves on
\citet{Skorstengaard+19b} by building those properties into the semantics of a
capability machine with a built-in call stack and call and return instructions,
which is proven fully abstract with respect to a more concrete capability
machine that replaces those pseudo-instructions with their calling convention.
Their proof of full abstraction uses a standard notion of components, which
import and export functions through their interfaces; their model of
observations is limited to cotermination.
\sna{Some of this seems like more detail than we need, and we need to
  compare ourselves to their list of criteria.}

>>>>

\paragraph{Formal Stack Safety on Capability Machines}
%
To our knowledge, the only line of work to date that has attempted
a positive and formal characterization of
stack safety is \citet{Skorstengaard+19b}, who introduce a calling
convention that uses local capabilities to preserve local state
encapsulation and well-bracketed control flow, using a logical
relation to reason about the stack safety of concrete
programs. Although the required hardware support is readily available
in capability machines like Cheri,
%% this technique incurs significant
%% costs, because
it requires the entire unused part of the stack to be
cleared whenever a security boundary is crossed. Their logical
relation captures capability safety without ``externally observable
side-effects (like console output or memory access traces)''
and can be used to reason about individual programs.
\citeauthor{Skorstengaard+19b} acknowledge the difficulty of defining
these concepts precisely:
``while [they] claim that [their] calling
convention enforces control-flow correctness, [they] do not prove a
general theorem that shows this, because it is not clear what such a
theorem should look like,'' noting that the correctness property
enjoyed by their technique ``is not made very explicit.''

StkTokens \citep{Skorstengaard+19} continues this work by defining a more efficient
calling convention that makes use of linear capabilities for stack and return
pointers.
%% The convention operates on a single shared stack and requires that
%% protected components avoid compromising their own security by following certain
%% simple rules---like not leaking their private capability seals. More
%% precisely but
Still informally, local state encapsulation is defined as
restricting accesses to the range of memory allocated to the current stack
frame, and well-bracketed control flow as only allowing returns from the topmost
frame to the immediately adjacent frame below. Formally, it improves on
\citet{Skorstengaard+19b} by building those properties into the semantics of a
capability machine with a built-in call stack and call and return instructions,
which is proven fully abstract with respect to a more concrete capability
machine.
%% that replaces those pseudo-instructions with their calling convention.
Their proof of full abstraction uses a standard notion of components, which
import and export functions through their interfaces.
%% ; their model of
%% observations is limited to cotermination.
\sna{Some of this seems like more detail than we need, and we need to
  compare ourselves to their list of criteria.}
\rb{Trimmed this a bit. Are you referring to their list of attack
  examples, or to something else?}
